Graphical Tests for Power Comparison of Competing Designs

### 2012-03-25 intro.tex
results

% !TEX root = comparison.tex
%% \section{Introduction} %for journal use above \firstsection{..} instead
%We introduce and discuss a tool for comparing different charts that are renderings of the same data in order to evaluate them for efficiency.
Graphics have a central role in the discovery process as well as in the presentation of results. 
As designer of  charts one of our main goals is to display data so that the information it contains is relayed in the most efficient way. 
In the search for the best display out of the seemingly endless multitude of methods we are guided by experience, knowledge of perceptual strengths and weaknesses, as well as aesthetic preferences. It is hard, if not impossible, to make the final decision on the design in an  objective manner. 
We can make this process less subjective by conducting user studies to evaluate competing designs of the same data, and measure designs for their efficiency by recording how accurately and quickly viewers can extract pieces of information. 
User studies for evaluating different designs has a long standing tradition. For  statistical graphics user studies have their origin in the (convenience) study on evaluating difficulty of a set of different visual tasks by \citet{cleveland:1984}, which has recently been repeated finding similar results based on a larger and more representative study by \citet{kosara:2010}.

@@

The strength of this evidence is determined by both the number of independent evaluations as well as the size of the lineup. (XXXX needs more motivation HH) The signal strength \cite{majumder:2011} of a particular lineup is given as:
\[
\text{signal strength} = 1- \left( \frac{x}{n}\right)^{1/(m-1)},
\]
where $m$ is the size of  the lineup, i.e. the number of individuals shown, and $x/n$ is the ratio of correctly identified data plots ($x$) in  $n$ independent evaluations. Signal strength is limited to a range of values between 0 and 1. Higher values indicate stronger signal.
 In the case of a parametric situation (i.e. the feature shown in the chart can be described in parametric form as a real-valued vector $\theta$ in the data), the signal strength is equivalent to the statistical power of the test.

>>>>

% !TEX root = comparison.tex
%% \section{Introduction} %for journal use above \firstsection{..} instead
%We introduce and discuss a tool for comparing different charts that are renderings of the same data in order to evaluate them for efficiency.
Graphics have a central role in the discovery process as well as in the presentation of results. 
As designer of  charts one of our main goals is to display data so that the information it contains is relayed in the most efficient way. 
In the search for the best display out of the seemingly endless multitude of methods we are guided by experience, knowledge of perceptual strengths and weaknesses, as well as aesthetic preferences. It is hard, if not impossible, to make the final decision on the design in an  objective manner. 
We can make this process less subjective by conducting user studies to evaluate competing designs of the same data, and measure designs for their efficiency by recording how accurately and quickly viewers can extract pieces of information. 
User studies for evaluating different designs has a long standing tradition. For  statistical graphics user studies have their origin in the (convenience) study on evaluating difficulty of a set of different visual tasks by \citet{cleveland:1984}, which has recently been replicated with similar results based on a large MTurk study by \citet{heer:2010}.

@@

The strength of this evidence is determined by both the number of independent evaluations as well as the size of the lineup. 
Assume a lineup of size $m$ is evaluated by $n$ independent observers, out of which $x$ identify the data plot correctly, which we can use as an estimate for the probability to correctly identify the data plot. Correctly identifying the data plot also depends on the features visible in the other plots. 
We can think of a lineup as $m-1$ head-to-head comparisons of the dataplot and $m-1$ nullplots, which the dataplot all has to win, in order to be picked as the plot with the most distinct features.  This introduces the signal strength \cite{majumder:2011} of a particular lineup as:
\[
\text{signal strength} = 1- \left( \frac{x}{n}\right)^{1/(m-1)},
\]
where $m$ is the size of  the lineup, i.e. the number of individuals shown, and $x/n$ is the ratio of correctly identified data plots ($x$) in  $n$ independent evaluations. Signal strength is limited to a range of values between 0 and 1. Higher values indicate stronger signal.
 In the case of a parametric situation (i.e. the feature shown in the chart can be described in parametric form as a real-valued vector $\theta$ in the data), the signal strength is equivalent to the statistical power of the test. While we can compute 95\% confidence intervals for signal strength by using 95\% confidence intervals for $x/n$ given as $x/n \pm 1.96 \cdot \sqrt{x/n \cdot (1-x/n)}$ limited to the interval [0,1], the resulting interval is generally to large to be useful.
